### Model calibration with ABC

Now we start to calibrate with approximate Bayesian computation. Our goal is to calibrate our SI vaccine trial model to the:  RV144 Thai vaccine trial; HVTN 702 South Africa vaccine trial; AMP trials of the VRC01 bnAb, and then use these calibrations to explore the parameter space that can lead to waning efficacy (within a trial) or different VE (between two trials).

In a regular epidemic modeling scenario (e.g. not a vaccine trial) we would be calibrating the parameters for which we don’t have exact/good empirical estimates; for this SI model that would include the risk multiplier and the subgroup frequencies of each risk group. But with this experiment one of our main goals is to identify a level of exposure heterogeneity in an HIV population that is consistent with some frailty VE results (e.g. either the ~0% VE in HVTN 702 or the waning PE in the AMP bnAB trials).  

Now, to do that we have to set target stats for the calibration. I first thought that we needed to set target stats for the placebo and vaccine arms separately. i.e. We first calibrate the model to the placebo arms, using incidence as the single target stat, and we vary only the `lambda` parameter. We get a value for `lambda` from the placebo calibrations that we then *set* for the vaccine calibrations, and in the vaccine calibrations we calibrate to the `risk multiplier` and the `initial conditions` (i.e., the %s of the population that are high and low risk, specifically.)

But, now I am thinking that we just do a rough first calibration of `lambda` to the placebo arm, then go straight to the parameter exploration on exposure heterogeneity and `epsilon` with VE as the target stat. The question then is, What parameter value combinations of `risk`, `high risk subgroup frequency`, and `epsilon` are consistent with some output value of VE.

### Target stat

So first specify the target stat.

```{r target stats}

time <- c(180,360,540,720,900,1080)  # every 6 months for 3 years
placebo.incidence <- rep(0.3, 6)    # flat incidence of 3.5% per 100 person years; adjust accordingly
VE.target <- rep(0.3, 6)
target.stats <- data.frame(time, placebo.incidence, VE.target)
```

### Function to get output VE for each simulation

```{r function for VE}

VE.from.sim <- function(mod){

mod <- mutate_epi(mod, total.Svh.Svm.Svl = Svh + Svm + Svl) #all susceptible in heterogeneous risk vaccine pop
mod <- mutate_epi(mod, total.Sph.Spm.Spl = Sph + Spm + Spl) #all susceptible in heterogeneous risk placebo pop

mod <- mutate_epi(mod, total.SIvh.SIvm.SIvl.flow = SIvh.flow + SIvm.flow + SIvl.flow) #all infections per day in heterogeneous risk vaccine pop
mod <- mutate_epi(mod, total.SIph.SIpm.SIpl.flow = SIph.flow + SIpm.flow + SIpl.flow) #all infections in heterogeneous risk placebo pop

mod <- mutate_epi(mod, rate.Vaccine.het = (total.SIvh.SIvm.SIvl.flow/total.Svh.Svm.Svl)*365*100)
mod <- mutate_epi(mod, rate.Placebo.het = (total.SIph.SIpm.SIpl.flow/total.Sph.Spm.Spl)*365*100)

mod <- mutate_epi(mod, VE.inst = 1 - rate.Vaccine.het/rate.Placebo.het)

return(mod)
}
```

### Calibration input model

Set up the function to wrap the EpiModel functions and make the incidence output file.  

```{r easyABC.input.model}

f <- function(x) {
  param <- param.dcm(beta = 0.004, 
                     c = 90/365,
                     prev = 0.10,    
                     #lambda = beta*c*prev,
                     #lambda = 0.000096,
                     lambda = 8e-06,
                     epsilon = x[1]) 
                     #risk = 10)
                     #risk = x[2])
  init <- init.dcm(Sp = 5000, Ip = 0,
                 Sv = 5000, Iv = 0,
                 Sph = 500, Iph = 0,    #placebo, high risk
                 Spm = 4000, Ipm = 0,   #placebo, medium risk
                 Spl = 500, Ipl = 0,    #placebo, low risk
                 Svh = 500, Ivh = 0,    #vaccine
                 Svm = 4000, Ivm = 0,   #vaccine
                 Svl = 500, Ivl = 0,    #vaccine
                 SIp.flow = 0, SIv.flow = 0, 
                 SIph.flow = 0, SIpm.flow = 0, SIpl.flow = 0,
                 SIvh.flow = 0, SIvm.flow = 0, SIvl.flow = 0)
  control <- control.dcm(nsteps = 365*3, new.mod = si_ode)
  mod <- dcm(param, init, control)
  
  #mod <- mutate_epi(mod, rate.Placebo = (SIp.flow/Sp)*365*100) #Incidence, placebo
  #mod <- mutate_epi(mod, rate.Placebo.het = (SIp.flow/Sp)*365*100) #Incidence, placebo
  
  #sim.ve <- VE.from.sim(mod)
  sim.ve <- mod.manipulate(mod)
  sim.ve <- as.data.frame(sim.ve)
  matchedTimes <- sim.ve$time %in% target.stats$time
  #out <- sim.ve$VE.inst[matchedTimes]
  out <- sim.ve$rate.Placebo.het[matchedTimes]
  return(out)
}
```

### Specify priors for parameters (or initial conditions)

The `beta` and `c` parameters are the transmission rate per contact and the contact rate, respectively. For now we keep these together in the `lambda` parameter, due to non-identifiability. Below we just set priors for `lambda`. We also will want to set priors for some of the initial conditions eventually. 

```{r priors}
# In order of "x" in the above function.
priors  <- list(#c("unif", 0.003, 0.008),      # beta
                #c("unif", 60/365, 120/365))    # c 
                #c("unif", 0.000005, 0.0001))  # lambda
                c("unif", 0.01, 0.5))         # epsilon
                #c("unif", 1, 20))            # risk multiplier
```


```{r eval=FALSE, include=FALSE}

fit.seq <- ABC_sequential(method = "Lenormand",
                       model = f,
                       prior = priors,
                       summary_stat_target = target.stats$placebo.incidence,
#                       summary_stat_target = target.stats$VE.target,
                       nb_simul = 10,
                       p_acc = 0.10,
                       alpha = 0.5,
                       progress_bar = TRUE)

```


```{r eval=FALSE, include=TRUE}

fit.rej <- ABC_rejection(model = f,
                     prior = priors,
                     nb_simul = 1000,
                     summary_stat_target = target.stats$placebo.incidence,
                     tol = 0.25,
                     progress_bar = TRUE)

```


```{r eval=FALSE, include=TRUE}

par(mfrow = c(1, 2))

fit <- fit.rej

plot(density(fit$param[, 1], from = 0.0000001,  to = 0.0001),
     #0.0000001, 0.00009
     main = "epsilon", 
     xlim = c(0.0000001, 0.0001),
     #ylim = c(0, 10),
     col=2)
#lines(density(fit$param[, 1], from = 0.01,  to = 0.5), col = 2)
abline(v = VE.target[1], lty = 2, col = 1)
legend("topright", legend = c("Truth", "Posterior"),
      lty = c(1, 2), col = 1:2, lwd = 2)

plot(density(fit$param[, 2], from = 1, to = 20),
     main = "risk multiplier", ylim = c(0, 0.5))
#lines(density(fit3$param[, 2], from = 1, to = 20), col = 2)
#abline(v = 1.5, lty = 2)
legend("topright", legend = c("Posterior"),
      lty = 1, col = 1, lwd = 2)
```

Other examples from Sam

### Use the mean of the parameters for model selection

```{r}

ms <- colMeans(fit$param)
param <- param.dcm(tau = ms[1], c = ms[2], D = 7)
init <- init.dcm(s.num = 9999, i.num = 1, si.flow = 0, is.flow = 0)
control <- control.dcm(nsteps = 104, new.mod = SISmod)
sim <- dcm(param, init, control)

par(mfrow = c(1,1))
plot(sim, y = "P")
arrows(myDat$time, myDat$uci, myDat$time, myDat$lci,
       col = "grey", len = 0.025, angle = 90, code = 3)
points(myDat$time, myDat$sampPrev, col = "black", pch = 16, cex = 1)

### Or use the full posterior distribution, as before

param <- param.dcm(tau = fit3$param[, 1], c = fit3$param[, 2], D = 7)
sim <- dcm(param, init, control)

plot(sim, y = "P")
arrows(myDat$time, myDat$uci, myDat$time, myDat$lci,
       col = "grey", len = 0.025, angle = 90, code = 3)
points(myDat$time, myDat$sampPrev, col = "black", pch = 16, cex = 1)
```
